### 2020-02-09 msr.tex
fixes and expand msr chapter

However, the successful demonstration of the \gls{ARE} spurred further
research into adapting \glspl{MSR} for civilian power generation
\cite{rosenthal_molten-salt_1970}. One of the key findings from the
research was that better economy could be achieved from breeding $^{233}$U
from $^{232}$Th in thermal-spectrum reactors than $^{239}$Pu from $^{238}$U
\cite{macpherson_molten_1985}. Ultimately, these efforts culminated in the
design, construction, and successful operation of the \gls{MSRE}, a graphite
moderated thermal \gls{MSR}. Although no breeding was attempted with the
\gls{MSRE}, scientists at \gls{ORNL} obtained a wealth of experimental data
and new insights from the study of this reactor. The \gls{MSRE} had a
graphite-moderated design with LiF-BeF$_2$-ZrF$_4$-UF$_4$ fuel salt mixture,
initially rated at 10 MW$_{\text{th}}$ but later restricted to 8 M
$_{\text{th}}$ due to a miscalculation of heat transfer capabilities
\cite{haubenreich_experience_1970}.

@@

Building on the \gls{MSRE}'s hugely successful run, \gls{ORNL} proposed a
major development project for the construction and operation of the
\gls{MSBR} \cite{macpherson_molten_1985}. However, they failed to secure
funding on two separate occasions in 1972 and 1974; they lost out to the
competing \gls{LMFBR} program which had a head start and wider
political and technical support. Nevertheless, from a technical perspective,
two independent technology evaluation and design studies of the \gls{MSR} had
``reported favorably on the promise of the system''
\cite{macpherson_molten_1985}.

>>>>

However, the successful demonstration of the \gls{ARE} spurred further
research into adapting \glspl{MSR} for civilian power generation
\cite{rosenthal_molten-salt_1970}. One of the key findings from the
research was that better economy could be achieved from breeding $^{233}$U
from $^{232}$Th in thermal-spectrum reactors than $^{239}$Pu from $^{238}$U
\cite{macpherson_molten_1985}. Ultimately, these efforts culminated in the
design, construction, and successful operation of the \gls{MSRE}, a graphite
moderated thermal \gls{MSR}. Although no breeding was attempted with the
\gls{MSRE}, scientists at \gls{ORNL} obtained a wealth of experimental data
and new insights from the study of this reactor. The \gls{MSRE} had a
graphite-moderated design with LiF-BeF$_2$-ZrF$_4$-UF$_4$ fuel salt mixture,
initially rated at 10 MW$_{\text{th}}$ but later restricted to 8
MW$_{\text{th}}$ due to a miscalculation of heat transfer capabilities
\cite{haubenreich_experience_1970}.

@@

Building on the \gls{MSRE}'s hugely successful run, \gls{ORNL} proposed a
new program for the construction and operation of a demonstration reactor
based on the \gls{MSBR} concept that they had developed
\cite{macpherson_molten_1985}. The \gls{MSBR} is a thermal-spectrum, single
fluid reactor with fertile $^{232}$Th isotopes mixed directly into the FLiBe
molten salt for $^{233}$U breeding \cite{gehin_liquid_2016}. Like the
\gls{MSRE}, the \gls{MSBR} relies on continuous online reprocessing to add
fertile material and remove fission product neutron poisons. As a breeder
reactor, the doubling time (the minimum amount of time required to produce
enough fissile material to start up another \gls{MSBR}) was estimated to be
approximately 22 years. However, \gls{ORNL} failed to secure funding on two
separate occasions in 1972 and 1974; they lost out to the competing
\gls{LMFBR} program which had a head start and wider political and technical
support. Nevertheless, from a technical perspective, two independent
technology evaluation and design studies of the \gls{MSR} had ``reported
favorably on the promise of the system'' \cite{macpherson_molten_1985}.

In spite of this setback, research into \glspl{MSR} continued through the late
1970s. In 1980, \gls{ORNL} published a report describing a new \gls{MSR}
concept, called the \gls{DMSR} \cite{gehin_liquid_2016}. This design was
developed in response to the fuel reprocessing restrictions introduced by
President Ford in 1976; the \gls{DMSR} is designed to operate as a
once-through converter system without fuel reprocessing. While it is largely
fueled by 19.75 \% \gls{HALEU}, the initial core loading includes thorium to
boost its conversion ratio throughout its lifetime. It has a continuous
online feed consists of \gls{HALEU} to maintain criticality, and denatured
$^{235}$U to keep uranium enrichement levels below nuclear non-proliferation
policy thresholds. The design also includes a gas sparging system for removing
gaseous \glspl{FP}, while noble metals naturally plate out onto the walls of
the coolant loop. A significant drawback in the \gls{MSBR} design is the
extensive neutron damage in the graphite moderator that necessitated frequent
replacement (every four years) throughout its operational lifetime. The
\gls{DMSR} avoids this issue by having a lower power density while maintaining
the overall power output of 2250 MW$_{\text{th}}$. As a result, the graphite
moderator was projected to last for the entirety of the \gls{DMSR}'s design
lifetime.

There was a concurrent program at the UK Atomic Energy Authority for the
development of a 2500MW$_{\text{e}}$ lead-cooled
\gls{MCFR} concept \cite{smith_assessment_1974}. It is a dual fluid system,
with separate loops for the fuel salt and the blanket salt. The blanket is a
1 m-wide tank surrounding the core. The relatively harder neutron spectrum
arising from the absence of moderators and the choice of chloride over
fluoride salt favors $^{239}$Pu breeding over the thorium cycle. Some
experiments were performed to study molten salt chemistry but no reactor
prototypes were built. The UK program was eventually shut down just like its
US counterpart partly due to the successful demonstration of the Prototype
Fast Reactor which had achieved criticality in 1974.

Following a lull lasting through the late 20th century, \gls{MSR} research
picked up pace due to renewed interest initiated by the \gls{GIF} in 2001.
Today, there are numerous \gls{MSR} concepts under active development led by
various national and commercial bodies. Many differences exist between
different concepts and the only common denominator is the use of nuclear fuel
dissolved in molten salt. Table * is a non-exhaustive list of \glspl{MSR}
under development.
